---
title: "Analysis of first 11 tumors"
author: "Mafe Senosain"
date: "1/24/2020"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/clustering_f.R")
```

```{r, echo=FALSE}
# Load data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/cellsubtypes.RData")
```

# All cells after denoising

## Major cell types

### Only major cell types
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$cell_type)

# Dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype3)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

```{r echo=FALSE, fig.width = 5, fig.height = 1.5}
par(mar=c(4,4,4,4))
par(font = 2) # bold axis
par(cex=1)
par(lwd = 1.2)
# Scatter plots
ggpubr::ggscatter(prcnt_by_pt, x = "Th_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF6BBC'),
          xlab = "CD4+ T cells", ylab = '')
ggpubr::ggscatter(prcnt_by_pt, x = "Tc_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF7500'),
          xlab = "CD8+ T cells", ylab = '')

epi_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/heatmap_epi_kmeans.png'
knitr::include_graphics(epi_path)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(annot_df, cell_type == 'immune')
sbst2 <- subset(annot_df, cell_type == 'endothelial')
sbst3 <- subset(annot_df, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(annot_df, cell_type == 'immune')
sbst2 <- subset(annot_df, cell_type == 'endothelial')
sbst3 <- subset(annot_df, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


# Exploring Rh negative only

```{r, echo=FALSE}
rd <- denoisingCTF::t_asinh(annot_df[,1:5])
k <- which(rd$`103Rh`< asinh(12/5))
rh_neg <- annot_df[k,]
```

## Major cell types

### Only major cell types
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$cell_type)

# Dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype3)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(rh_neg, cell_type == 'immune')
sbst2 <- subset(rh_neg, cell_type == 'endothelial')
sbst3 <- subset(rh_neg, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(rh_neg, cell_type == 'immune')
sbst2 <- subset(rh_neg, cell_type == 'endothelial')
sbst3 <- subset(rh_neg, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


# Exploring Rh positive only
```{r, echo=FALSE}
rh_pos <- annot_df[-k,]
```

## Major cell types

### Only major cell types
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$cell_type)

# Dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype3)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(rh_pos, cell_type == 'immune')
sbst2 <- subset(rh_pos, cell_type == 'endothelial')
sbst3 <- subset(rh_pos, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(rh_pos, cell_type == 'immune')
sbst2 <- subset(rh_pos, cell_type == 'endothelial')
sbst3 <- subset(rh_pos, cell_type == 'mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


# Cell types afected by removal of Rh+
```{r, echo=FALSE}
cat('Total number of cells: ', nrow(annot_df), '\n')
cat('Rh Negative number of cells: ', nrow(rh_neg), '\n')
cat('Rh Positive number of cells: ', nrow(rh_pos), '\n')
cat('% of cells presumably lost in previous analysis: ', nrow(rh_pos)/nrow(annot_df)*100, '%', '\n')

```

## Total cell types percentages:
```{r, echo=FALSE}
table(annot_df$cell_type)/nrow(annot_df)*100
table(annot_df$subtype)/nrow(annot_df)*100
table(annot_df$subtype2)/nrow(annot_df)*100
table(annot_df$subtype3)/nrow(annot_df)*100
```

## Rh negative cell types percentages:
```{r, echo=FALSE}
table(rh_neg$cell_type)/nrow(rh_neg)*100
table(rh_neg$subtype)/nrow(rh_neg)*100
table(rh_neg$subtype2)/nrow(rh_neg)*100
table(rh_neg$subtype3)/nrow(rh_neg)*100
```

## Rh positive cell types percentages:
```{r, echo=FALSE}
table(rh_pos$cell_type)/nrow(rh_pos)*100
table(rh_pos$subtype)/nrow(rh_pos)*100
table(rh_pos$subtype2)/nrow(rh_pos)*100
table(rh_pos$subtype3)/nrow(rh_pos)*100
```


```{r, echo=FALSE}
library(png)
subtypes_all_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_all.png'
subtypes_rhneg_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_rhneg.png'
subtypes_rhnpos_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_rhpos.png'
```

## Heatmaps
### All cells
```{r, echo=FALSE}
knitr::include_graphics(subtypes_all_path)
```

### Rh negative
```{r, echo=FALSE}
knitr::include_graphics(subtypes_rhneg_path)
```

### Rh positive
```{r, echo=FALSE}
knitr::include_graphics(subtypes_rhnpos_path)
```

