---
title: "Analysis of first 11 tumors"
author: "Mafe Senosain"
date: "1/24/2020"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/30_DA_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/40_DE_functions.R")
```

```{r}
ClusterEval_plot <- function(data, data_type = c('heatmap', 'tSNE', 'UMAP'), cluster_col, ft_cols){
    # Heatmap
    if(data_type == 'heatmap'){
        library(gplots)
        scale_max = max(data)
        heat_palette_med <- colorRampPalette(c("black", "yellow","#FAF7C9"))
        pairs.breaks_med <- c(seq(0, scale_max/6.6, by = 0.1),
                              seq(scale_max/6.6, scale_max/3.3, by = 0.1),
                              seq(scale_max/3.3, scale_max, by = 0.1))

        hm_pl <- heatmap.2(data,
                  main = "Median protein expression",
                  dendrogram = "both",
                  Rowv = TRUE,
                  Colv = TRUE,
                  breaks = pairs.breaks_med,
                  revC = FALSE,
                  symkey = FALSE,
                  symbreaks = FALSE,
                  scale = "none",
                  cexRow = 0.8,
                  cexCol = 0.8,
                  key = TRUE,
                  col = heat_palette_med,
                  trace = "none",
                  density.info = 'none',
                  sepcolor="#424242",
                  margins = c(6,10),
                  colsep=1:ncol(data),
                  rowsep=1:nrow(data),
                  sepwidth=c(0.005,0.005),
                  keysize = 1,
                  key.title = 'Intensity',
                  key.xlab= "Arcsinh Transform",
                  extrafun = box(lty = "solid"),
                  cellnote = data,
                  notecol = 'red',
                  srtCol=45
                  )
        return(hm_pl)
    }

    if(data_type == 'tSNE'){
        # Plotting
        library(ggplot2)
        library(gridExtra)
        library(reshape2)
        data <- data[c(colnames(data)[ft_cols], cluster_col, 'tSNE1', 'tSNE2')]
        test <- melt(data, id= c(cluster_col, 'tSNE1', 'tSNE2'))
        var_list <- unique(test$variable)
        pl <- list()

        # Defining cluster centers
        edata <- cbind(data$tSNE1, data$tSNE2, cluster_col=data[cluster_col])
        colnames(edata) <- c('x', "y", "z")
        center <- aggregate(cbind(x,y) ~ z, data = edata, median)

        p_cl <- ggplot(data, aes_string(x='tSNE1', y='tSNE2', color = data[,cluster_col]))+
          geom_point(alpha=0.6, shape = 19) + theme_bw() + ggtitle(cluster_col)+
          annotate("text", label = center[,1], x=center[,2], y = center[,3],
                   size = 6, colour = "black", fontface = 'bold')
        for(i in seq_along(var_list)) {
          p <- ggplot(subset(test,test$variable==var_list[i]), aes(x=tSNE1, y=tSNE2, colour = value)) +
            geom_point(alpha=0.4) + theme_bw() + ggtitle(var_list[i]) +
            scale_colour_gradient(low = "gray75", high = "blue")
          pl[[i]] <- p
        }

        plot(p_cl)
        tsne_pl <- grid.arrange(grobs=pl)
    }

    if(data_type == 'UMAP'){
        # Plotting
        library(ggplot2)
        library(gridExtra)
        library(reshape2)
        data <- data[c(colnames(data)[ft_cols], cluster_col, 'UMAP1', 'UMAP2')]
        test <- melt(data, id= c(cluster_col, 'UMAP1', 'UMAP2'))
        var_list <- unique(test$variable)
        pl <- list()

        # Defining cluster centers
        edata <- cbind(data$UMAP1, data$UMAP2, cluster_col=data[cluster_col])
        colnames(edata) <- c('x', "y", "z")
        center <- aggregate(cbind(x,y) ~ z, data = edata, median)

        p_cl <- ggplot(data, aes_string(x='UMAP1', y='UMAP2', color = data[,cluster_col]))+
          geom_point(alpha=0.6, shape = 19) + theme_bw() + ggtitle(cluster_col)+
          annotate("text", label = center[,1], x=center[,2], y = center[,3],
                   size = 6, colour = "black", fontface = 'bold')
        for(i in seq_along(var_list)) {
          p <- ggplot(subset(test,test$variable==var_list[i]), aes(x=UMAP1, y=UMAP2, colour = value)) +
            geom_point(alpha=0.4) + theme_bw() + ggtitle(var_list[i]) +
            scale_colour_gradient(low = "gray75", high = "blue")
          pl[[i]] <- p
        }
        
        plot(p_cl)
        umap_pl <- grid.arrange(grobs=pl)
    }
}
```

```{r, echo=FALSE}
# Load data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/cellsubtypes.RData")
```

```{r}
annot_df <- annot_df[-which(annot_df$CANARY =='I'),]
ref <- ref[-which(ref$CANARY =='I'),]

# Changing un-identified T cells into "Other_immune"
k <- which(annot_df$subtype =='T_cells')
k2 <- which(annot_df$subtype =='NK_cells')

annot_df['subtype'] <- as.character(annot_df$subtype)
annot_df[c(k,k2), 'subtype'] <- 'Other_immune'
annot_df$subtype <- as.factor(annot_df$subtype)

annot_df['subtype4'] <- as.character(annot_df$subtype)
annot_df[which(annot_df$cell_type=='Epithelial'), 'subtype4'] <- 'Epithelial'
annot_df$subtype4 <- as.factor(annot_df$subtype4)
```


```{r}
dendrogram_barplot <- function(data, dist_method = 'euclidean', 
    hclust_method = 'ward.D', corr_mat = FALSE){
    
    # Dendrogram
    if(corr_mat){
        corr_pt <- Hmisc::rcorr(t(as.matrix(data)), type = 'spearman')
        dist_mat <- dist(corr_pt$r, method = dist_method)
    } else {
        dist_mat <- dist(data, method = dist_method)
    }
    hclust_avg <- hclust(dist_mat, method = hclust_method)
    par(mar=c(2,7,4,2), lwd=2, mgp=c(3,1,0), las=1, font.lab=2)
    plot(hclust_avg,cex = 0.8, hang = -1, main = paste0(hclust_method, ' linkage'))
    
    # Barplot
    coul = brewer.pal(9, "Set1")
    data <- data[hclust_avg$order,] #dendrogram order
    data <- t(as.matrix(data))
    
    par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7.5), xpd=TRUE, font.lab=2, lwd=1) # axis label locations
    barplot(data,
            col=coul ,
            border='white',
            horiz=TRUE,
            cex.names=0.75,
            cex.axis = 0.75)
    title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
    title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
    legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.3,0),
               cex = 0.75)

}

```

# All cells after denoising

## Major cell types

### Only major cell types

```{r}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$cell_type)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

# Correlation plot
corr_plot(prcnt_by_pt)

ggpubr::ggscatter(prcnt_by_pt, x = "Endothelial", y = 'Immune',
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'gray'),
          xlab = "Endothelial cells", ylab = 'Immune cells')

# Differential abundance
x <- DA_analysis(annot_df, ref, class_col='cell_type', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals

```



## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D', corr_mat = T)
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete', corr_mat = T)

# Correlation plot
corr_plot(prcnt_by_pt)

# Differential abundance
x <- DA_analysis(annot_df, ref, class_col='subtype', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals
```

```{r echo=FALSE, fig.width = 5, fig.height = 2}
# Scatter plots
ggpubr::ggscatter(prcnt_by_pt, x = "Th_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF6BBC'),
          xlab = "CD4+ T cells", ylab = '')
ggpubr::ggscatter(prcnt_by_pt, x = "Tc_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF7500'),
          xlab = "CD8+ T cells", ylab = '')
    
epi_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/heatmap_epi_kmeans.png'
knitr::include_graphics(epi_path)
```

### Subtypes (Epithelial cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype4')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype4)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D', corr_mat = T)
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete', corr_mat = T)

# Correlation plot
corr_plot(prcnt_by_pt)

# Differential abundance
x <- DA_analysis(annot_df, ref, class_col='subtype4', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'Immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

# Correlation plot
corr_plot(prcnt_by_pt)

# Differential abundance
x <- DA_analysis(sbst, ref, class_col='subtype', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals

```


## Only Epithelial

### All Epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'Epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D', corr_mat = T)
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete', corr_mat = T)

# Correlation plot
corr_plot(prcnt_by_pt)

# Differential abundance
x <- DA_analysis(sbst, ref, class_col='subtype', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(annot_df, cell_type == 'Immune')
sbst2 <- subset(annot_df, cell_type == 'Endothelial')
sbst3 <- subset(annot_df, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms and barplots
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D')
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete')

dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'ward.D', corr_mat = T)
dendrogram_barplot(prcnt_by_pt, dist_method = 'euclidean', hclust_method = 'complete', corr_mat = T)

# Correlation plot
corr_plot(prcnt_by_pt)

# Differential abundance
x <- DA_analysis(sbst, ref, class_col='subtype', group_levels=c('G','P'),
     group_col = 'CANARY', ptID_col = 'pt_ID')
x$pvals
```



## Differential protein expression

```{r}
med_endo <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='cell_type',
    celltype_name='Endothelial', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_epi <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='cell_type',
    celltype_name='Epithelial', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_fibmes <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='cell_type',
    celltype_name='Mesenchymal', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_imm <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='cell_type',
    celltype_name='Immune', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_CD8T <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='subtype',
    celltype_name='Tc_cells', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_CD4T <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='subtype',
    celltype_name='Th_cells', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

med_mye <- median_by_pt(annot_df, ref, subset_celltype=T, celltype_col='subtype',
    celltype_name='Myeloid', ask_features=F, ft_idxs = c(15, 17:31, 33:35, 37:48, 50, 51), 
    ptID_col = 'pt_ID', groups=c('G', 'P'))

```


## Endothelial cells
```{r echo=FALSE}
sbst <- reshape::melt(med_endo$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_endo$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```


## Fibroblasts/Mesenchymal cells
```{r echo=FALSE}
sbst <- reshape::melt(med_fibmes$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_fibmes$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```


## Immune cells
```{r echo=FALSE}
sbst <- reshape::melt(med_imm$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_imm$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```




## CD8 T cells
```{r echo=FALSE}
sbst <- reshape::melt(med_CD8T$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_CD8T$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```


## CD4 T cells
```{r echo=FALSE}
sbst <- reshape::melt(med_CD4T$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_CD4T$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```


## Myeloid cells
```{r echo=FALSE}
sbst <- reshape::melt(med_mye$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_mye$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```



## Epithelial cells
```{r echo=FALSE}
sbst <- reshape::melt(med_epi$Med_expression)
#plot(ftimp_rf)
ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,10)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22))

med_epi$pvals
```

```{r echo=FALSE}
plotly::ggplotly(ggplot(sbst, aes(x=CANARY, y=value, color = CANARY)) +
  geom_boxplot() +
  ylim(0,6)+
  ggsignif::geom_signif(comparisons = list(c("G", "P")), 
       map_signif_level=TRUE) +
  facet_wrap(~variable) +
  theme(plot.title = element_text(hjust = 0.5, size=22)))
```



# Cell type proyection

```{r}
# sampling k cells per sample
smpl <- annot_df[1,]
smpl <- smpl[-1,]
k = 2000
pts = unique(annot_df$pt_ID)
for(i in pts){
  #print(dim(annot_df[which(annot_df$pt_ID == i),]))
  x <- dplyr::sample_n(annot_df[which(annot_df$pt_ID %in% i),], size=k, replace=F)
  smpl <- rbind(smpl, x)
}
smpl <- denoisingCTF::t_asinh(smpl)
```

```{r}
#print(as.matrix(colnames(annot_df)))
#col_id <- c(15, 17:31, 33:35, 37, 38, 40:51)
col_id <- c(15, 19,20, 29:31, 34,37,38, 40,41, 44, 46,47, 50)

#col_id <- c(15, 20,29:31,34,37,40,44,46,47)


# tSNE
tsne_smp <- Rtsne::Rtsne(smpl[,col_id],  check_duplicates = FALSE)
tsne_smp <- as.data.frame(cbind(smpl,
                                    tSNE1 = tsne_smp$Y[,1],
                                    tSNE2 = tsne_smp$Y[,2]))
# UMAP
umap_smp <- umap::umap(smpl[,col_id])
umap_smp <- as.data.frame(cbind(smpl,
                                    UMAP1 = umap_smp$layout[,1],
                                    UMAP2 = umap_smp$layout[,2]))
```



```{r}
ClusterEval_plot(tsne_smp, data_type = 'tSNE', cluster_col= 'subtype4', ft_cols = col_id)
ClusterEval_plot(umap_smp, data_type = 'UMAP', cluster_col= 'subtype4', ft_cols = col_id)

```

