---
title: "Analysis of first 11 tumors"
author: "Mafe Senosain"
date: "1/24/2020"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
```

```{r, echo=FALSE}
# Load data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/cellsubtypes.RData")
```

```{r}
annot_df <- annot_df[-which(annot_df$CANARY =='I'),]
ref <- ref[-which(ref$CANARY =='I'),]
```

```{r}
corr_pt <- Hmisc::rcorr(t(as.matrix(prcnt_by_pt)), type = 'spearman')
dist_mat <- dist(corr_pt$r)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1)
```
# All cells after denoising

## Major cell types

### Only major cell types

```{r}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$cell_type)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)

```



### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype3)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

```{r echo=FALSE, fig.width = 5, fig.height = 2}
# Scatter plots
ggpubr::ggscatter(prcnt_by_pt, x = "Th_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF6BBC'),
          xlab = "CD4+ T cells", ylab = '')
ggpubr::ggscatter(prcnt_by_pt, x = "Tc_cells", y = c('Epithelial_7', 'Epithelial_8', 'Epithelial_9'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = '#FF7500'),
          xlab = "CD8+ T cells", ylab = '')
    
epi_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/heatmap_epi_kmeans.png'
knitr::include_graphics(epi_path)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(annot_df$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'Immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'Immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All Epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type == 'Epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(annot_df, cell_type == 'Immune')
sbst2 <- subset(annot_df, cell_type == 'Endothelial')
sbst3 <- subset(annot_df, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(annot_df, cell_type == 'Immune')
sbst2 <- subset(annot_df, cell_type == 'Endothelial')
sbst3 <- subset(annot_df, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


# Exploring Rh negative only

```{r, echo=FALSE}
rd <- denoisingCTF::t_asinh(annot_df[,1:5])
k <- which(rd$`103Rh`< asinh(12/5))
rh_neg <- annot_df[k,]
```

## Major cell types

### Only major cell types
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$cell_type)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype3)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_neg, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_neg$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'Immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'Immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All Epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(rh_neg, cell_type == 'Epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(rh_neg, cell_type == 'Immune')
sbst2 <- subset(rh_neg, cell_type == 'Endothelial')
sbst3 <- subset(rh_neg, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(rh_neg, cell_type == 'Immune')
sbst2 <- subset(rh_neg, cell_type == 'Endothelial')
sbst3 <- subset(rh_neg, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Dendrograms
dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat, method = 'ward.D')
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1, main = 'Ward Linkage')

dist_mat <- dist(prcnt_by_pt)
hclust_avg <- hclust(dist_mat)
par(mar=c(2,7,4,2), lwd=2)
plot(hclust_avg,cex = 0.8, hang = -1,  main = 'Complete Linkage')

# Barplot
coul = brewer.pal(9, "Set1")
data <- prcnt_by_pt[hclust_avg$order,] #dendrogram order
data <- t(as.matrix(data))

par(mgp=c(1.5,0.25,0), las=1, mar = c(3, 5, 1, 7), xpd=TRUE, font.lab=2) # axis label locations
barplot(data,
        col=coul ,
        border='white',
        horiz=TRUE,
        cex.names=0.75,
        cex.axis = 0.75)
title(xlab="Cell type (% of sample)", mgp = c(1.5, 0.5, 0))
title(ylab="Patient ID",  mgp = c(3, 0.1, 0))
legend('topright', legend = rownames(data), fill = coul, ncol = 1, inset=c(-0.25,0),
           cex = 0.75)

# Correlation plot
corr_plot(prcnt_by_pt)
```


# Exploring Rh positive only
```{r, echo=FALSE}
rh_pos <- annot_df[-k,]
```

## Major cell types

### Only major cell types
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'cell_type')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$cell_type)

# Dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes and T cells as a group
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype3')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype3)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes

### All subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=rh_pos, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(rh_pos$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Immune

### All subtypes
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'Immune')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)

```

### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'Immune')
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


## Only Epithelial

### All Epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(rh_pos, cell_type == 'Epithelial')
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst1 <- subset(rh_pos, cell_type == 'Immune')
sbst2 <- subset(rh_pos, cell_type == 'Endothelial')
sbst3 <- subset(rh_pos, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype <- factor(sbst$subtype)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


### Subtypes (T cells as one group)
```{r, echo=FALSE}
sbst1 <- subset(rh_pos, cell_type == 'Immune')
sbst2 <- subset(rh_pos, cell_type == 'Endothelial')
sbst3 <- subset(rh_pos, cell_type == 'Mesenchymal')
sbst <- rbind(sbst1, sbst2, sbst3)
sbst$subtype2 <- factor(sbst$subtype2)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype2')
rownames(prcnt_by_pt) <- paste0(rownames(prcnt_by_pt), '_', ref$CANARY)

levels(sbst$subtype2)

# Cluster dendrogram and barplot
dendrogram_barplot(prcnt_by_pt)
# Correlation plot
corr_plot(prcnt_by_pt)
```


# Cell types afected by removal of Rh+
```{r, echo=FALSE}
cat('Total number of cells: ', nrow(annot_df), '\n')
cat('Rh Negative number of cells: ', nrow(rh_neg), '\n')
cat('Rh Positive number of cells: ', nrow(rh_pos), '\n')
cat('% of cells presumably lost in previous analysis: ', nrow(rh_pos)/nrow(annot_df)*100, '%', '\n')

```

## Total cell types percentages:
```{r, echo=FALSE}
table(annot_df$cell_type)/nrow(annot_df)*100
table(annot_df$subtype)/nrow(annot_df)*100
table(annot_df$subtype2)/nrow(annot_df)*100
table(annot_df$subtype3)/nrow(annot_df)*100
```

## Rh negative cell types percentages:
```{r, echo=FALSE}
table(rh_neg$cell_type)/nrow(rh_neg)*100
table(rh_neg$subtype)/nrow(rh_neg)*100
table(rh_neg$subtype2)/nrow(rh_neg)*100
table(rh_neg$subtype3)/nrow(rh_neg)*100
```

## Rh positive cell types percentages:
```{r, echo=FALSE}
table(rh_pos$cell_type)/nrow(rh_pos)*100
table(rh_pos$subtype)/nrow(rh_pos)*100
table(rh_pos$subtype2)/nrow(rh_pos)*100
table(rh_pos$subtype3)/nrow(rh_pos)*100
```


```{r, echo=FALSE}
library(png)
subtypes_all_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_all.png'
subtypes_rhneg_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_rhneg.png'
subtypes_rhnpos_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/woCD90/figs/rh_comparison/subtypes_rhpos.png'
```

## Heatmaps
### All cells
```{r, echo=FALSE}
knitr::include_graphics(subtypes_all_path)
```

### Rh negative
```{r, echo=FALSE}
knitr::include_graphics(subtypes_rhneg_path)
```

### Rh positive
```{r, echo=FALSE}
knitr::include_graphics(subtypes_rhnpos_path)
```

# Other Comparisons

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'cell_type', cell_pop = 'Endothelial')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'cell_type', cell_pop = 'Mesenchymal')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'cell_type', cell_pop = 'Immune')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'subtype', cell_pop = 'Myeloid')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'subtype2', cell_pop = 'T_cells')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'subtype', cell_pop = 'NK_cells')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'subtype', cell_pop = 'Th_cells')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'subtype', cell_pop = 'Tc_cells')
```

```{r, echo=FALSE, fig.width = 7, fig.height = 7, warning=FALSE}
boxplots_m(annot_df, col_name = 'cell_type', cell_pop = 'Epithelial')
```
