---
title: "Summary of CyTOF dataset"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
library(jcolors)
library(ComplexHeatmap)
knitr::opts_chunk$set(echo = FALSE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/30_DA_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/40_DE_functions.R")
```

```{r, echo=FALSE}
# Load data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/cellsubtypes.RData")
# read edited CSV file as CDE_TMA36
CDE_TMA36 <- read.csv(file = '/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2020FEB25_SA.csv')
```

```{r, echo=FALSE}
# Data pre-processing
# Changing un-identified T cells into "Other_immune"
k <- which(annot_df$subtype_B =='T_cells')
k2 <- which(annot_df$subtype_B =='NK_cells')

annot_df['subtype_B'] <- as.character(annot_df$subtype_B)
annot_df[c(k,k2), 'subtype_B'] <- 'Other_immune'
annot_df$subtype_B <- as.factor(annot_df$subtype_B)

annot_df['subtype_B4'] <- as.character(annot_df$subtype_B)
annot_df[which(annot_df$cell_type_B=='Epithelial'), 'subtype_B4'] <- 'Epithelial'
annot_df$subtype_B4 <- as.factor(annot_df$subtype_B4)
annot_df$subtype_B4 <- factor(annot_df$subtype_B4, levels = c("Epithelial", "Endothelial", "Fib_Mesenchymal", "Other_immune", "Myeloid", "Tc_cells", "Th_cells"))

# All cell/sub types B have Fib_Mesenchymal
ref <- ref[-42,] # sample 13376 has been collected twice and will be merged, only one ref is needed
grp <- 'B'
CDE_TMA36 <- CDE_TMA36[match(ref$pt_ID, CDE_TMA36$pt_ID),] # Select only pts with CyTOF data

```

# Abundance analysis

## Major cell types

### Major cell types (immune as one)
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = paste0('cell_type_', grp))
```

```{r, echo=FALSE}
# Dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)

Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score", 
  heatmap_legend_param = list(color_bar = "continuous"), 
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  top_annotation = ha)

# Correlation plot
corr_plot(prcnt_by_pt)
```


### Major cell types and immune subtypes
```{r, echo=FALSE}
prcnt_by_pt <- ClassAbundanceByPt(data=annot_df, ptID_col = 'pt_ID', 
    class_col = paste0('subtype_', grp, '4'))
prcnt_by_pt['ALL_Immune'] <- rowSums(prcnt_by_pt[,4:7])
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
set.seed(101)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha,
        column_km = 2)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Subtypes
### All subtypes

```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A != 'Epithelial')
sbst[,paste0('subtype_', grp)] <- factor(sbst[,paste0('subtype_', grp)])
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = paste0('subtype_', grp))
```

```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A == 'Epithelial')
sbst$subtype_A <- factor(sbst$subtype_A)
prcnt_by_ptE <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype_A')
prcnt_by_pt <- cbind(prcnt_by_pt/2, prcnt_by_ptE/2)
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha)

# Correlation plot
corr_plot(prcnt_by_pt)
```

#### Correlation of Epithelial HLA-DR+ subsets and immune cells
```{r echo=FALSE, fig.width = 5, fig.height = 2}
epi_path <- '/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/figs/heatmap_epi_kmeans.png'
knitr::include_graphics(epi_path)
```

```{r echo=FALSE, fig.width = 12, fig.height = 4}
# Scatter plots
ggpubr::ggscatter(prcnt_by_pt*100, x = "Myeloid", y = c('Epithelial_8', 'Epithelial_9', 'Epithelial_10'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
          xlab = "Myeloid cells", ylab = 'Relative abundance (%)')
ggpubr::ggscatter(prcnt_by_pt*100, x = "Tc_cells", y = c('Epithelial_8', 'Epithelial_9', 'Epithelial_10'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
          xlab = "CD8+ T cells", ylab = 'Relative abundance (%)')

ggpubr::ggscatter(prcnt_by_pt*100, x = "Th_cells", y = c('Epithelial_8','Epithelial_9', 'Epithelial_10'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
          xlab = "CD4+ T cells", ylab = 'Relative abundance (%)')
    
```

### All subtypes (epi clusters grouped)
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A != 'Epithelial')
sbst[,paste0('subtype_', grp)] <- factor(sbst[,paste0('subtype_', grp)])
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = paste0('subtype_', grp))
```

```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A == 'Epithelial')
sbst$subtype_A <- factor(sbst$subtype_A)
prcnt_by_ptE <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype_A')

prcnt_by_ptE <- cbind(Epi_123=rowSums(prcnt_by_ptE[,c(1,3,4)]), Epi_456=rowSums(prcnt_by_ptE[,5:7]), 
                     Epi_7=prcnt_by_ptE$Epithelial_7, Epi_8910=rowSums(prcnt_by_ptE[,c(2,9,10)]))

prcnt_by_pt <- cbind(prcnt_by_pt/2, prcnt_by_ptE/2)
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
set.seed(101)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha,
        column_km = 3)

# Correlation plot
corr_plot(prcnt_by_pt)
```

```{r}
load("/Users/senosam/Documents/Massion_lab/RNASeq_summary/rnaseq.RData")
clust_eigen <- data.frame(read_tsv('/Users/senosam/Documents/Massion_lab/RNASeq_summary/Results_23_Sep_20_2/Eigengenes.tsv'))
rownames(clust_eigen) <- clust_eigen$X1
clust_eigen$X1 <- NULL
clust_eigen <- t(clust_eigen)

p_all <- p_all[na.omit(match(p_all$Vantage_ID, rownames(clust_eigen))),]
p_all <- p_all[match(rownames(prcnt_by_pt), p_all$pt_ID),]
#p_all <- p_all[na.omit(match(p_all$pt_ID, rownames(prcnt_by_pt))),]
clust_eigen <- data.frame(clust_eigen[match(p_all$Vantage_ID, rownames(clust_eigen)),])
rownames(clust_eigen) <- rownames(prcnt_by_pt)
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    C0=clust_eigen$C0,
    C1=clust_eigen$C1,
    C2=clust_eigen$C2,
    C3=clust_eigen$C3,
    simple_anno_size = unit(0.5, "cm")
)
set.seed(50)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha)


```

#### Correlation of Epithelial HLA-DR+ subsets and immune cells
```{r echo=FALSE, fig.width = 12, fig.height = 4}

# Scatter plots
ggpubr::ggscatter(prcnt_by_pt*100, x = 'Epi_8910', y = c("Myeloid", 'Tc_cells', 'Th_cells'),
          add = "reg.line", conf.int = TRUE, combine = TRUE,
          cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
          xlab = "Epi_8910", ylab = 'Relative abundance (%)')

```

## Epithelial cells only

### All Epithelial subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A == 'Epithelial')
sbst$subtype_A <- factor(sbst$subtype_A)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype_A')
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha)

# Correlation plot
corr_plot(prcnt_by_pt)
```

### All Epithelial subtypes (combined clusters)
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A == 'Epithelial')
sbst$subtype_A <- factor(sbst$subtype_A)
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = 'subtype_A')

prcnt_by_pt <- cbind(Epi_123=rowSums(prcnt_by_pt[,c(1,3,4)]), Epi_456=rowSums(prcnt_by_pt[,5:7]), 
                     Epi_7=prcnt_by_pt$Epithelial_7, Epi_8910=rowSums(prcnt_by_pt[,c(2,9,10)]))
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
set.seed(101)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha,
        column_km = 2)

# Correlation plot
corr_plot(prcnt_by_pt)
```

## Stromal cells only

### All subtypes
```{r, echo=FALSE}
sbst <- subset(annot_df, cell_type_A != 'Epithelial')
sbst[,paste0('subtype_', grp)] <- factor(sbst[,paste0('subtype_', grp)])
prcnt_by_pt <- ClassAbundanceByPt(data=sbst, ptID_col = 'pt_ID', 
    class_col = paste0('subtype_', grp))
```

```{r, echo=FALSE}
# Cluster dendrogram and barplot
par(mar=c(3.5,6,1,2))
par(mgp=c(3,0.5,0))
dendrogram_barplot(prcnt_by_pt)

ha = HeatmapAnnotation(
    CANARY = as.factor(CDE_TMA36$CANARY),
    simple_anno_size = unit(0.5, "cm")
)
Heatmap(t(as.matrix(scale(prcnt_by_pt))), name = "z-score",
        heatmap_legend_param = list(color_bar = "continuous"), 
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        top_annotation = ha)

# Correlation plot
corr_plot(prcnt_by_pt)
```

