---
title: "TMA36 CyTOF dataset: Indolent vs. Aggressive"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
library(jcolors)
library(ComplexHeatmap)
knitr::opts_chunk$set(echo = FALSE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/51_supervised_analysis_viz.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
```

```{r, echo=FALSE}
# Load data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/cellclusters.RData")
ref <- ref[-42,] # sample 13376 has been collected twice and will be merged, only one ref is needed
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/percent_pt.RData")
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/protein_correlations.RData")
```

```{r, echo=FALSE}
# Clinical data
CDE <- read.csv(file = '/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2021JAN13_SA_MF.csv')
```

```{r, echo=FALSE}
CDE <- CDE[match(ref$pt_ID, CDE$pt_ID),] # Select only pts with CyTOF data
```

```{r, echo=FALSE}
# Add new class and SILA to annot_df
pts = unique(annot_df$pt_ID)
annot_df['SILA'] <- NA
annot_df['n_op1'] <- NA
for (i in pts) {
    k <- which(annot_df$pt_ID == i)
    annot_df[k, 'SILA'] <- CDE[which(CDE$pt_ID == i), 'SILA']
    annot_df[k, 'n_op1'] <- CDE[which(CDE$pt_ID == i), 'n_op1']
}
```
# 1. Entire data set

## 1.1 UMAPs 
```{r}
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_allcelltypes.RData")
```

```{r, echo=FALSE}
# By density, cluster, patient
#plot_colors
m_ct <- c("#FA0000", "#207BBB", "#00B84C", "#999999")
s_ct <- c("#f0dc78", "#FA0000", "#207BBB", "#00B84C", "#631461", "#966995", "#c9c9c9", "#FF7500" ,"#FF6BBC")
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
#pie(rep(1,n), col=sample(col_vector, n))


ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'cell_type_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = m_ct, lg_names = levels(umap_smp$cell_type_B),  x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'subtype_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = s_ct, lg_names = levels(as.factor(umap_smp$subtype_B)),  x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),  x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop2,  x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),  x_lim = x_lim, y_lim = y_lim)

```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
x <- c(15,20,29:31,34,37,40,44,46,47,50)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = x, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### UMAP of Indolent cells only

```{r, echo=FALSE}
k <- which(umap_smp$n_op1 == 'ind')
umap_smp_i <- umap_smp[k,]
ClusterUMAP_plot(umap_smp_i, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_i, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'cell_type_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = m_ct, lg_names = levels(umap_smp_i$cell_type_B),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_i, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'subtype_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = s_ct, lg_names = levels(as.factor(umap_smp_i$subtype_B)),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_i, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp_i$pt_ID)),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_i, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = c('#3498DB'), lg_names = c('indolent'),  x_lim = c(-11,11), y_lim = c(-11,11))

```

### UMAP of Aggressive cells only

```{r, echo=FALSE}
k <- which(umap_smp$n_op1 == 'agg')
umap_smp_a <- umap_smp[k,]
ClusterUMAP_plot(umap_smp_a, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_a, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'cell_type_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = m_ct, lg_names = levels(umap_smp_a$cell_type_B),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_a, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'subtype_B', plot_clusnames = F, plot_title = 'Cell identity',
                 plot_colors = s_ct, lg_names = levels(as.factor(umap_smp_a$subtype_B)),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_a, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp_a$pt_ID)),  x_lim = c(-11,11), y_lim = c(-11,11))
ClusterUMAP_plot(umap_smp_a, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = c('aggressive'),  x_lim = c(-11,11), y_lim = c(-11,11))
```

## 1.2 Barplots of subtypes clusters
### Epithelial cancer cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Epithelial', ptID_col = 'pt_ID', celltype_nm_plot = "Epithelial cancer cells", plotly_p = T)
```

### Endothelial cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Endothelial', ptID_col = 'pt_ID', celltype_nm_plot = "Endothelial", plotly_p = T)
```

### Fibroblasts/Mesenchymal cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Fib_Mesenchymal', ptID_col = 'pt_ID', celltype_nm_plot = "Fibroblasts/Mesenchymal cells", plotly_p = T)
```

### CD4+ T cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Th_cells', ptID_col = 'pt_ID', celltype_nm_plot = "CD4+ T cells", plotly_p = T)
```

### CD8+ T cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Tc_cells', ptID_col = 'pt_ID', celltype_nm_plot = "CD8+ T cells", plotly_p = T)
```

### DN T cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='DNT_cells', ptID_col = 'pt_ID', celltype_nm_plot = "DN T cells", plotly_p = T)
```

### Myeloid cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='Myeloid', ptID_col = 'pt_ID', celltype_nm_plot = "Myeloid cells", plotly_p = T)
```

### NK cells
```{r, echo=FALSE}
main_bp(annot_df, cond_col='n_op1', celltype_col='subtype_B', cluster_col='clusters_B', celltype_nm='NK_cells', ptID_col = 'pt_ID', celltype_nm_plot = "NK cells", plotly_p = T)
```


## 1.3 Heatmaps by cell type proportion

### 1.3.1 Main cell types

```{r, echo=FALSE, fig.width = 10, fig.height = 3}
frac_hm(prcnt_celltypes, CDE, 'n_op1')
```

### 1.3.2 All subtypes

```{r, echo=FALSE, fig.width = 10, fig.height = 3}
frac_hm(prcnt_subtypes, CDE, 'n_op1')
```

### 1.3.3 Stroma + immune subtypes

```{r, echo=FALSE, fig.width = 10, fig.height = 3}
rc_prcnt <- function(dt, col_rm){
  dt[,col_rm]<- NULL
  tt <- apply(dt, 1, sum)
  dt <- dt/tt
}

x <- rc_prcnt(prcnt_subtypes, col_rm = 'Epithelial')
frac_hm(x, CDE, 'n_op1')
```

### 1.3.4 Immune subtypes

```{r, echo=FALSE, fig.width = 10, fig.height = 3}
x <- rc_prcnt(prcnt_subtypes, col_rm = c('Epithelial', 'Endothelial', 'Fib_Mesenchymal'))
frac_hm(x, CDE, 'n_op1')
```

## 1.4 Boxplots comparing cell type proportion
### 1.4.1 Main cell types

```{r, echo=FALSE}
frac_boxplot(prcnt_celltypes, CDE, 'n_op1')
```

### 1.4.2 All subtypes

```{r, echo=FALSE}
frac_boxplot(prcnt_subtypes, CDE, 'n_op1')
```

### 1.4.3 All cell type clusters

```{r, echo=FALSE, fig.width = 15, fig.height = 10}
frac_boxplot(prcnt_clusters, CDE, 'n_op1')
```

# 2. In depth analysis by cell subtype

## 2.1 Epithelial cancer cells (ECC)
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_epi.RData")
# subset index
st_nm <- 'Epithelial'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(15,18,19,21:23,26,28,29,31,33,35,37,38,40,41,50,51)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(1:8)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```


## 2.2 Endothelial cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_endo.RData")
# subset index
st_nm <- 'Endothelial'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18, 19, 21, 22, 26, 28, 31, 33, 35, 38, 41, 42, 50, 51)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(14:16)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```


## 2.3 Fibroblasts/Mesenchymal cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_fmes.RData")
# subset index
st_nm <- 'Fib_Mesenchymal'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18, 19, 21, 22, 26, 27, 28, 33, 35, 38, 41, 42, 43, 48, 50, 51)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(9:13)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```

## 2.4 CD8+ T cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_cd8t.RData")
# subset index
st_nm <- 'Tc_cells'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18:22, 26, 28, 31, 35, 42, 50)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(21:23)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```


## 2.5 CD4+ T cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_cd4t.RData")
# subset index
st_nm <- 'Th_cells'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18:22, 26, 28, 31, 35, 42, 50)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(17:20)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```


## 2.6 DN T cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_dnt.RData")
# subset index
st_nm <- 'DNT_cells'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18:22, 26, 28, 31, 35, 42, 50)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(24:27)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```

## 2.7 Myeloid cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_mye.RData")
# subset index
st_nm <- 'Myeloid'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18:22, 26, 28, 31, 35, 42, 50)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(28:30)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```


## 2.8 NK cells
```{r, echo=FALSE}
# load UMAP data
load("/Users/senosam/Documents/Massion_lab/CyTOF_summary/both/analysis/supervised/UMAP_nk.RData")
# subset index
st_nm <- 'NK_cells'
k <- which(annot_df$subtype_B == st_nm)
# ft cols
ft_cols <- c(18:22, 26, 28, 31, 35, 42, 50)
```

### UMAPs
```{r}
nop1 <- c("#EC7063", "#3498DB")
nop2 <- c("#EC7063", "#3498DB", "grey72")

x_lim <- c(min(umap_smp$UMAP1)-1, max(umap_smp$UMAP1)+1)
y_lim <- c(min(umap_smp$UMAP2)-1, max(umap_smp$UMAP2)+1)

#x and y lim based on UMAP1 and UMAP2 max and min
library(scales)
set.seed(101)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
clus_col = hue_pal()(length(unique(umap_smp$clusters_B)))

ClusterUMAP_plot(umap_smp, density = T, color_by_cluster = F, color_by_protein = F, pallete = F,
                 plot_clusnames = F, plot_title = 'Density', x_lim = x_lim, y_lim = y_lim)

ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'clusters_B', plot_clusnames = F, plot_title = 'Cluster identity',
                 plot_colors = clus_col, lg_names = levels(umap_smp$clusters_B),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F,
                 cluster_col = 'n_op1', plot_clusnames = F, plot_title = 'Behavior class',
                 plot_colors = nop1, lg_names = levels(umap_smp$n_op1),   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = F, color_by_continuous = T, pallete = F,
                 cluster_col = 'SILA', plot_clusnames = F, plot_title = 'SILA score',
                 plot_colors = nop1,   x_lim = x_lim, y_lim = y_lim)
ClusterUMAP_plot(umap_smp, color_by_cluster = T, color_by_protein = F, pallete = F, plot_colors = col_vector,
                 cluster_col = 'pt_ID', plot_clusnames = F, plot_title = 'Patient ID',
                 lg_names = levels(as.factor(umap_smp$pt_ID)),   x_lim = x_lim, y_lim = y_lim)
```

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
ClusterUMAP_plot(umap_smp, color_by_cluster = F, color_by_protein = T, 
                 ft_cols = ft_cols, plot_clusnames = F, x_lim = x_lim, y_lim = y_lim, 
                 nrow_prot = 2, heights_prot = unit(c(2,2), c("in", "in")))
```

### Heatmap (median exp per cluster)
```{r, echo=FALSE}
df_cluster <- denoisingCTF::t_asinh(cbind(annot_df[k,ft_cols], 
                                          'cluster'= annot_df[k,'clusters_B'], 
                                          'n_op1'=annot_df[k,'n_op1']))
```

```{r, echo=FALSE, fig.width = 10, fig.height = 6}
hm_median(df_cluster)
```

### Histograms (protein expression per cluster + distribution)
```{r, echo=FALSE, fig.width = 9, fig.height = 9}
hist_prot(df_cluster)
```

### Heatmap of proportions
```{r, echo=FALSE, fig.width = 10, fig.height = 3}
col_rm <- colnames(prcnt_clusters)[-c(31:34)]
x <- rc_prcnt(prcnt_clusters, col_rm = col_rm)
frac_hm(x, CDE, 'n_op1')
```

### Boxplots comparing cluster proportions (% of sample)
```{r, echo=FALSE, fig.width = 10, fig.height = 6}
frac_boxplot(x, CDE, 'n_op1')
```



# 3. Other visualizations

## 3.1 Protein-protein correlation (co-expression)
### Bulk
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_bulk)
```

### Epithelial cancer cells

```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_epi)
```

### Endothelial cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_endo)
```

### Fibroblasts/Mesenchymal cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_fmes)
```

### CD4 T cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_th)
```

### CD8 T cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_tc)
```

### DN T cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_dnt)
```

### Myeloid cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_mye)
```

### NK cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_nk)
```

### Other immune cells
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
protein_corr(co_oimm)
```

## 3.2 Cell type proportions correlations
### Major cell types
```{r, echo=FALSE}
corr_plot(prcnt_celltypes)
```

### Cell subtypes
```{r, echo=FALSE}
corr_plot(prcnt_subtypes)
```

### Cell subtypes clusters
```{r, echo=FALSE, fig.width = 15, fig.height = 10}
corr_plot(prcnt_clusters)
```

